---
title: "Title"
author: "Put your name here!"
output: 
  html_document:
    toc: true
    toc_float: true
    df_print: paged
    code_download: true
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```

```{r libraries}
library(tidyverse)         # for graphing and data cleaning
```

```{r data}
data(mpg) # this dataset is part of an R package
```

Just like I did above, at the top of the file, put three R code chunks. The first controls options. Right now I have it commented out (a `#` sign in front), but you should uncomment (remove the `#`) when your document is complete so you don't see messages and warnings. The second loads libraries. It is nice to give a brief description of what the library does to remind yourself why you are loading it. The third loads any data used in the document.

The yaml section of the document (the part at the very top, between the set of three dashes) is EXTREMELY picky but it also gives you a way to customize your document. If you want to learn more about changing this section see [section 3.1.4](https://bookdown.org/yihui/rmarkdown/html-document.html#appearance-and-style) of the *R Markdown: The Definitive Guide* by Yihui Xie, J. J. Allaire, and Garrett Grolemund.

Use headings and subheadings to number your problems, like I've done below.  Leave a space before and after each R code chunk, so as not to mess up the numbering. Write in complete sentences to answer your questions. Your comments should clearly reference the R code and output that you include. Do not include R code and output that you don't reference.

## Exercise 1

Description of exercise 1 ...

### a.  

Description of exercise 1a. 

If necessary, explain the code that will be run in the next section (a really easy example, in this case). The following chunk of code adds 2 plus 3.

```{r}
2 + 3
```

Afterwords, explain what the results show. For example, as we can see, 2 + 3 is 5. 

### b.

## Exercise 2

### a. 

### b. 

...



