---
title: "tidy practice with Billboard 100"
output: 
  html_document:
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: false
    df_print: paged
    code_download: true
---

## Advanced data wrangling

## dplyr and tidyr working together

Watch the following video that succinctly goes through some examples of data wrangling using the `lubridate`, `stringr` and `tidyr` packages, as well as more familiar code from `dplyr`. You may want to watch several times, or slow the playback speed down!

You can then download the `Rmd` file (using the `Code` button in the top right corner) to check out the code. 

<iframe width="560" height="315" src="https://www.youtube.com/embed/DiY8EqZDwoI" title="YouTube video player" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

```{r}
library(dplyr)
library(readr)
billboard100 <- read_csv("data/billboard100.csv")
```
```{r}
music_df <-  billboard100 %>%
  select(date, rank:artist, weeks_popular='weeks-on-board')
music_df
```

```{r}
library(lubridate)
library(stringr)
```


```{r}
music_df.cleaned <-  
  music_df %>%
  mutate(date = ymd(date),
         primary_artist = ifelse(str_detect(artist, 'Featuring'),
                                 str_match(artist, '(.*)\\sFeaturing')[,2],
                                 artist),
         featured_artist = str_match(artist, 'Featuring\\s(.*)')[,2])
```

```{r}
library(tidyr)
```

```{r}
music_df.cleaned %>%
  distinct(song, primary_artist, featured_artist) %>%
  pivot_longer(2:3, names_to="artist_type", values_to="artist_name") %>%
  filter(artist_name=="Drake") %>%
  count(artist_type)
```

```{r}
music_df.cleaned %>%
  filter(rank <= 3) %>%
  select(date, rank, song) %>%
  pivot_wider(names_from="rank",values_from="song")
```


```{r}
# Load mtv data to capture genre

 mtv_data <-  read_csv("https://gist.githubusercontent.com/mbejda/9912f7a366c62c1f296c/raw/dd94a25492b3062f4ca0dc2bb2cdf23fec0896ea/10000-MTV-Music-Artists-page-1.csv")

glimpse(mtv_data)
```

```{r}
top5_genres = music_df.cleaned %>%
  inner_join(mtv_data, by = c("primary_artist" = "name")) %>%
  select(song, primary_artist, genre) %>%
  filter(!is.na(genre)) %>%
  distinct() %>%
  count(genre) %>%
  top_n(5) %>%
  pull(genre)
```

```{r}
library(ggplot2)
```

```{r}
music_df.cleaned %>%
  inner_join(mtv_data, by = c("primary_artist" = "name")) %>%
  mutate(date=floor_date(date, unit="year")) %>%
  select(date, song, genre) %>%
  filter(genre %in% top5_genres) %>%
  count(date, genre) %>%
  ggplot(aes(x=date,y=n,color=genre)) + 
  geom_line()
```




