---
title: "Project Data Proposal"
author: "WRITE YOUR NAME HERE"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(tidyverse)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 42 was an arbitrary one on my part.
set.seed(42)
```

# Big-picture

## Research question

*What is your research question?*

## Description of data

*Please give a very short description of the data set along with it's original source.*

## Load data into R

*Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680). If the data shouldn't be published online, then please submit the spreadsheet file on Moodle.*

```{r, message=FALSE}

```

## Clean variable names

*Piping your data frame into the `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with.*

```{r}

```

## Explore your data

*Be sure to explore your data. Note that `eval=FALSE` is set so that R Markdown doesn't "evaluate" this code chunk, i.e. it will ignore it in the ultimate `.html` report. You should run this code on your own, but not in the ultimate `.html` report.*

```{r, eval = FALSE}
glimpse()
```

------------------------------------------------------------------------

# Variables

## Identification variable

*What is your identification (ID) variable (if you have one)?*

## Outcome variable

*What is your outcome variable* $y$? What are its units of measurement?

## Numerical explantory variable

*What is your numeric explanatory variable? What are its units of measurement?*

## Categorical explantory variable

*What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels.*

------------------------------------------------------------------------

# Rows/observations

## Observational units

*What is the observational unit of your data? In other words, what does each row in your data represent?*

## Sample size

*How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?*

------------------------------------------------------------------------

# Preview of data

## Pare down variables

*`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at.*

1.  *The identification variable*
2.  *The outcome variable* $y$
3.  *The numerical explanatory variable*
4.  *The categorical explanatory variable*
5.  *Optional: any other variable you find interesting*

```{r}

```

## Preview data

*Display a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package . You'll get the same 5 rows everytime you knit this document and hence replicable results because we set the seed value of the random number generator in the first code chunk above.*

```{r}

```
