---
title: "Your ISS Data Project Title"
author: "WRITE YOUR NAME HERE"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    fig_caption: yes
    theme: lumen
    toc: yes
    toc_depth: 2
    df_print: kable
    toc_float:
      collapsed: no
---

```{r, message=FALSE}
# Load all packages here:
library(tidyverse)
library(janitor)

# Set seed value of random number generator to get "replicable" random numbers.
# The choice of seed value of 42 was an arbitrary one on my part.
set.seed(42)

# read_gdoc_csv function for sharing data on Google sheets
read_gdoc_csv <- function(link) {
  sheet_id <- stringr::str_extract(link, "(?<=d/)[^/]+")
  tab_id <- stringr::str_extract(link, "(?<=gid=)[0-9]+")
  csv <- stringr::str_glue("https://docs.google.com/spreadsheets/d/{sheet_id}/export?format=csv&gid={tab_id}")
  
  tryCatch({
    suppressWarnings({
      readr::read_csv(csv)
    })
  }, error = function(e) {
    if (grepl("401", e$message) || grepl("403", e$message) || grepl("permission", e$message, ignore.case = TRUE)) {
      message("\nERROR: Unable to access the Google Sheet (HTTP error 401/403)")
      message("This typically means the sheet isn't publicly accessible.")
      
      message("\nTO FIX THIS:")
      message("If you OWN this sheet:")
      message("1. Open the Google Sheet in your browser")
      message("2. Click 'Share' in the top-right corner")
      message("3. Click 'Change' next to 'Restricted'")
      message("4. Select 'Anyone with the link'")
      message("5. Set permission to 'Viewer' (not 'Editor')")
      message("6. Click 'Done'")
      
      message("\nIf you DO NOT OWN this sheet:")
      message("You'll need to ask the owner to complete the steps above to make it public.")
      message("Politely request they share it with 'Anyone with the link' as a 'Viewer'.")
      
      message("\nIMPORTANT NOTE: The URL should look like this when shared correctly:")
      message(paste0("https://docs.google.com/spreadsheets/d/", sheet_id, "/edit?gid=", tab_id))
      
      message("\nAfter permissions are updated, wait a minute and try again.")
    } else {
      message("Error: ", e$message)
    }
    return(invisible(NULL))
  })
}
```

# Big-picture

## Research question

*What is your research question?*

## Description of data

*Please give a very short description of the data set along with it's original source.*

## Load data into R

*Include the code to load your data here. If your data is not confidential nor is it private in nature, consider publishing it as a `.csv` file on Google Sheets as in the code chunk below; instructions on how to do this are in Steps 1-6 [here](https://twitter.com/rudeboybert/status/1055821846539591680), or simply use the customised read_gdoc_csv() function.*

```{r, message=FALSE}

```

## Clean variable names

*Piping your data frame into the `clean_names()` function from the `janitor` package will clean your variable names, making them easier to work with.*

```{r}

```

## Explore your data

*Be sure to explore your data. Note that `eval=FALSE` is set so that R Markdown doesn't "evaluate" this code chunk, i.e. it will ignore it in the ultimate `.html` report. You should run this code on your own, but not in the ultimate `.html` report.*

```{r, eval = FALSE}
glimpse()
```
*Write any observations about your exploration here.*
------------------------------------------------------------------------

# Variables

## Identification variable

*What is your identification (ID) variable (if you have one)?*

## Outcome variable

*What is your outcome variable* $y$? What are its units of measurement?

## Numerical explantory variable

*What is your numeric explanatory variable? What are its units of measurement?*

## Categorical explantory variable

*What is your categorical explanatory variable? Does it have between 3 and 5 levels. Please list the different levels.*

------------------------------------------------------------------------

# Rows/observations

## Observational units

*What is the observational unit of your data? In other words, what does each row in your data represent?*

## Sample size

*How many rows/cases are in the data i.e. what is the sample size? Is the sample size at least 50?*

------------------------------------------------------------------------

# Preview of data

## Pare down variables

*`select()` the following variables **in this order** and drop all others. Eliminating all unnecessary variables will making visually exploring the raw values less taxing mentally, as we'll have less data to look at.*

1.  *The identification variable*
2.  *The outcome variable* $y$
3.  *The numerical explanatory variable*
4.  *The categorical explanatory variable*
5.  *Optional: any other variable you find interesting*

```{r}

```

## Preview data

*Display a random sample of 5 rows of your data frame by piping it into the `sample_n(5)` function from the `dplyr` package . You'll get the same 5 rows everytime you knit this document and hence replicable results because we set the seed value of the random number generator in the first code chunk above.*

```{r}

```


# Exploratory data analysis

_Let's do an little exploratory data analysis._

## Inspect for missing values

_Address missing values._

Using the trick in the "Become an R Data Ninja!" file, we get a sense of how many missing values we have in our data. 

```{r}
colSums(is.na(ma_schools))
```

Missing values details...

```{r}

```

## Summary statistics

_Compute some quick summary statistics of the outcome variable and comment._

```{r}

```

Summary statistics details...

## Histogram of outcome variable 

_Visualize the distribution of the outcome variable using a histogram and comment._

```{r, fig.cap = "Figure 1. Detialed title", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot() 
```

These data seem roughly bell-shaped, with no obvious skew. There is an outlier average Math SAT score of around 750 or so. This school is the [MA Academy for Math and Science School](https://www.massacademy.org/).


## Scatterplot 

_Visualize the relationship of the outcome variable and the numerical explanatory variable using a scatterplot and comment._

```{r, fig.cap = "Figure 2. Detialed title", fig.align = "center", fig.width=16/2, fig.height=9/2}

ggplot()
```

Some visualization details...


## Boxplot

_Visualize the relationship of the outcome variable and the categorical explanatory variable using a scatterplot and comment._

```{r, fig.cap = "Figure 3. Detialed title", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot()
```

Some visualization details...


## Colored scatterplot

_Visualize the relationship of the outcome variable and both explanatory variables using a colored scatterplot and comment._

```{r, fig.cap = "Figure 4. Detialed title", fig.align = "center", fig.width=16/2, fig.height=9/2}
ggplot()
```

Some visualization details

# 3. Multiple linear regression

## 3.1 Methods

The components of our multiple linear regression model are the following:

* Outcome variable $y$ =
* Numerical explanatory variable $x_1$ = 
* Categorical explanatory variable $x_2$ = 


## 3.2 Model Results

***

Table 2. Regresstion table of...
 
```{r}

```

***

## 3.3 Interpreting the regression table

The regression equation for ... is the following:

$$
\begin{aligned}\widehat {score} =& b_{0} + b_{prct} \cdot prct + b_{med} \cdot 1_{is\ med}(x_2) + b_{large} \cdot 1_{is\ large}(x_2) \\
 =& 588.190 - 2.777 \cdot prct - 11.913 \cdot 1_{is\ med}(x_2) - 6.362 \cdot 1_{is\ large}(x_2)
\end{aligned} 
$$

* The intercept ($b_0$ = 588.190) represents .... (Table 2).
* The estimate for the slope  ($b_{prct}$ = -2.777) is the .... Based on this estimate, for every .., there was an associted increase/decrease in ...
* The estimate for ... ($b_{med}$ = -11.913) and ... ($b_{large}$ = -6.36) are the offsets in intercept relative to the baseline group's　... (Table 2). In other words, on average ...

Thus the three regression lines have equations:

$$
\begin{aligned}
\text{Small schools (in red)}: \widehat {score} =& 588.190 - 2.777 \cdot prct\\
\text{Medium schools (in green)}: \widehat {score} =& 576.277 - 2.777 \cdot prct\\
\text{Large schools (in blue)}: \widehat {score} =& 581.828 - 2.777 \cdot prct
\end{aligned} 
$$


***


# 4. Discussion 

## 4.1 Conclusions


## 4.2 Limitations


## 4.3 Further questions

